Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-05-14 paper.tex
appendix

\section{Detailed modelling of starlight and dust in clusters}
\label{sect:SD}
In this section we present the details of our model of stars and dust
(SD) in clusters of galaxies. In the first part we the measured
spectral distribution of light from SD in a typical galaxy to derive a
simple spectral model for SD in galaxy clusters. This is done by
renormalizing the galactic SD light to match the observed luminosity
from a cluster in the far infrared (IR) and ultraviolet (UV) bands. In
the second part we derive the spatial distribution of the SD light
from a stacked sample of clusters.

\subsection{Energy spectrum}
The SD light from clusters is dominated by the emission from
galaxies. This radiation is expected to leak out into the much larger
volume of the intra-cluster medium, hence we approximate the spectral
distribution of the intra-cluster light (ICL) with the light from
galaxies. We use the far IR to UV spectrum derived in reference
\cite{2006ApJ...648L..29P} for a Milky Way (MW) type galaxy to
characterize the spectral distribution of SD in a cluster. Note that
we only keep the spectral shape, and renormalize the spectral
distribution using the luminosity from SD in clusters.

In Fig.~\ref{fig:SD_spectra} we show the spectral distribution of the
light from SD together with the cosmic microwave background (CMB)
black-body distribution in a $6\times10^{14}\,\msun$ galaxy
cluster. We fit the spectral shape of SD individually; the dust that
peaks at about $10^{-2}\,\ev$ is fitted using a double power-law,
while the broader spectral distribution of the stars that peaks at
about $1\,\ev$ is fitted using a tipple power-law. The best fit to the
galactic spectrum is given by:
\begin{eqnarray}
  u_\stars^\gal(\eph) &=& \frac{23\,\rmn{eV}}{\rmn{cm}^3} 
  \left(\frac{1.23\,\rmn{eV}}{\eph}\right)^{1.9} \nonumber \\
  &\times&\left[1+\left(\frac{2.04\,\rmn{eV}}{\eph}\right)^{20}\right]
  ^{-\frac{1.9}{20}}\nonumber \\
  &\times& \left[1+\left(\frac{0.78\,\rmn{eV}}{\eph}\right)^{20}\right]^{-\frac{2.6}{20}}\,, \\
  u_\dust^\gal(\eph) &=& 
  \frac{40\,\rmn{eV}}{\rmn{cm}^3} 
  \left(\frac{0.0144\,\rmn{eV}}{\eph}\right)^{4.9}\nonumber \\
  &\times& \left[1+\left(\frac{0.0144\,\rmn{eV}}{\eph}\right)^{1.9}\right]^{-4.9}\,.
\end{eqnarray}
The specific energy density of the light from SD in a cluster can now
be derived from
\begin{eqnarray} 
u_\sd(\eph, r) &=& j(r)\,u_\sd(\eph)\,,
\label{eq:u_SD_er}
\end{eqnarray}
where the spectral distribution of SD is given by
\begin{eqnarray}
  u_\sd(\eph) &\equiv& \eph^2\frac{\dd^2 N_\ph}{\dd \eph \dd V}
  =  \sum_i N_i(\mvir) u_i^\gal(\eph)\,,\nonumber \\ 
\rmn{for}\quad i&=&\{\rmn{stars,dust}\}\,.
\end{eqnarray}
Here $\eph$ denotes the energy of the SD photons, $j(r)$ describes the
unitless spatial distribution of SD (derived in
sect.~\ref{sect:SD_radial}), and $N_i(\mvir)$ denotes the mass
dependent normalization for the SD. This function is determined using
the total energy of the light from SD within $\rvir$ which is
represented by $E_{i,\vir}$. We relate this energy to the luminosity
$L_i$ of each respective wavelength band (optical and IR light for the
dust, and UV light for the stars) using
\begin{eqnarray} 
  E_{i,\vir} &=& L_i \frac{\rvir}{c} \nonumber \\
  &=&N_i(\mvir)\int_{\rvir} \int_i \,j(r) 
  \frac{u_i^\gal(\eph)}{\eph}\,\dd V\dd \eph\,,\nonumber \\
 \rmn{for}\quad i&=&\{\rmn{stars,dust}\}\,.
\label{eq:E_SD}
\end{eqnarray}
Here we approximate the total energy of the photons within $\rvir$
with the luminosity \cite{2008A&A...490..547G} multiplied the typical
timescale it takes for a photon to propagate through the cluster (we
assume that the cluster is optically thin).

We derive the {\em luminosity from the stars} using
\begin{eqnarray}
L_\stars(M_{200\rmn{av}})&=&10^{\frac{m_\odot-m_\rmn{cl}-2.5\log\left(1+z\right)}{2.5}}
\left(\frac{D_\clu}{D_\odot}\right)^2 L_\odot\nonumber\\
&\approx& 5.3\times10^{44}\,\rmn{erg}\,\rmn{s}^{-1}\,,
\label{eq:L_stars}
\end{eqnarray}
where the average apparent magnitude of a cluster at $z\approx 0.25$
in the $(r+i)$-band is given by $m_\rmn{cl}\approx 15.5$
\cite{2005MNRAS.358..949Z}. This magnitude is derived from stacked
clusters observed by the SDSS with the average mass
$M_{200\rmn{av}}=4.0\times10^{13}\msun$. We also use an apparent
magnitude of the sun $m_\odot\approx -27.7$, distance to the sun
$D_\odot=4.85\times10^{-6}\,\rmn{pc}$, luminosity of the sun
$L_\odot=1.17\times10^{33}\,\rmn{erg\,s}^{-1}$, and distance to the
cluster $D_\clu=1.26\times10^9\,\rmn{pc}$. To account for different
cluster masses, we attach a simple mass scaling to the luminosity from
the stars in clusters in Eq.~(\ref{eq:L_stars}).  Since a large
contribution of the total starlight in a cluster comes from the
brightest cluster galaxy (BCG), we assume that the total starlight has
the same halo mass scaling as the BCG \cite{2010ApJ...713.1037H}
\begin{equation}
L_\stars(\mvir)=L_\stars(M_{200\rmn{av}})\,
\left(\frac{\mvir}{M_{200\rmn{av}}}\right)^{0.18}\,.
\label{eq:L_stars_m}
\end{equation} 
We can now fix the unitless normalization constant for the stars in
Eq.~(\ref{eq:E_SD}) by integrating over the cluster volume and
spectral distribution of the stars:
\begin{equation}
 N_\stars(\mvir) = 
\left(\frac{\mvir}{M_{200\rmn{av}}}\right)^{0.18}\,
\frac{6.0\,\rvir\times10^{-9}\,\kpc^2}{\int_{\rvir} j(r) \,\dd V}\,.
\label{eq:N_stars}
\end{equation}

@@

\subsection{Radial profiles}
\label{sect:SD_radial}
Our goal in this section is to derive a simple spatial model for the
distribution of the light from SD in galaxy clusters. For this reason
we use stacked cluster observations by Sloan in the $r$ and $i$-band
that trace the stars and for simplicity we assume that the dust trace
the stars in the clusters {\bf Christoph, why is this a valid
  approximation?}. The stacked surface brightness profiles in
\cite{2005MNRAS.358..949Z} are measured at redshift $z \approx 0.25$
with an average mass of the clusters of $4.0\times10^{13}\,\msun$. The
profiles are build up out of three components; diffuse ICL, galaxies,
and the BCG in the center of clusters. However, the total volume
overlap of the inter-stellar medium with a cluster is very small
compared to the total overlap of the ICL- and BCG with a cluster, thus
the relative contribution of the galaxies to the IC emission is
suppressed (c.f. Eq.~\ref{eq:SD_overlap}). We remove this bias in our
benchmark model for SD by cutting out the galaxies which roughly
corresponds to $70$ percent of the total SD light. In
Fig.~\ref{fig:SD_spatial} we show the SDSS stacked brightness profiles
as well as the fitted profiles\footnote{The measured brightness is
  converted into units of $\rmn{erg}\,\rmn{s}^{-1}$ using
  \cite{2010...book} $S(\rmn{mag}\,''^{-2}) =
  \mathcal{M}_\odot+21.6-2.5\log_{10}[S(L_\odot\,\rmn{pc}^{-2})]$,
  where the sun's absolute magnitude in the $r$-band is given by
  $\mathcal{M}_\odot=27.1$ \cite{1998gaas.book.....B} and the
  luminosity of the sun by $L_\odot=3.85\times10^{33}\,
  \rmn{erg}\,\rmn{s}^{-1}$.}, where both the data and fits are
normalized with the central surface brightness of the total SD
component ($S_\sd^\rmn{tot}(0)$). Our benchmark spatial model is shown
with the solid black line, and include the contribution from the ICL
and BCG.

@@

Instead of modelling the surface brightness with a de Vaucouleur
profile with the addition of a power-law, we use a double beta profile
model for simplicity of deprojection. It is given by
\begin{equation}
\tilde{S}_\sd (r_\bot) = \frac{S_\sd (r_\bot)}{S_\rmn{SD}^\rmn{tot}(0)} = 
\sum_{i=1}^2 \,\tilde{S}_i\, 
\left[1 + \left( \frac{r_\bot}{r_{\mathrm{c}_i}}\right)^2\right]
^{-3\beta_i + 1/2}\,,
\label{double_beta}
\end{equation}
where we found the following best fitted normalized central brightness
($S_i$), core radius ($r_{\mathrm{c}_i}$), and slope ($\beta_i$) for
the total model:
\begin{align}
&\tilde{S}_1^\rmn{tot} = 1.0\,,
&&r_{\mathrm{c}_1}^\rmn{tot} = 1.8\,\rmn{kpc}\,,
&&&\beta_1^\rmn{tot} = 0.45\, \nonumber\\
&\tilde{S}_2^\rmn{tot} = 2.3\times10^{-3}\,,
&&r_{\mathrm{c}_2}^\rmn{tot} = 0.19\,\rvir\,,
&&&\beta_2^\rmn{tot} = 0.44\,.\nonumber\\
& && &&&
\label{fit_spatial_IR}
\end{align}
For the smooth model, which is also our benchmark SD model, we find:
\begin{align}
&\tilde{S}_1^\rmn{sm}=0.47,\,
&&r_{\mathrm{c}_1}^\rmn{sm}=3.9\,\rmn{kpc}\,,
&&&\beta_1^\rmn{sm}=0.53 \,\nonumber\\
&\tilde{S}_2^\rmn{sm}=8.3\times10^{-4}\,,
&&r_{\mathrm{c}_2}^\rmn{sm}=0.19\,\rvir\,,
&&&\beta_2^\rmn{sm}=0.54\,.\nonumber\\
& && &&&
\label{fit_spatial_IR_sm}
\end{align}
The three dimensional spatial profile is derived by deprojecting the
surface brightness in Eq.~(\ref{double_beta}) (see
e.g. \cite{2004A&A...413...17P} for details about the deprojection):
\begin{eqnarray}
  j(r)  = \sum_{i=1}^2 \frac{\tilde{S}_i}{2\pi\,r_{\mathrm{c}_i}}\,
  \frac{6 \beta_i - 1}{\left(1 + r^2/r^2_{\mathrm{c}_i}\right)^{3\beta_i}}\,
  \mathcal{B}\left(\frac{1}{2},3\beta_i\right)\,,\nonumber\\
\end{eqnarray}
where $\mathcal{B}(a,b)$ denotes the beta-function
\cite{1965hmfw.book.....A}.

The spatial distribution of the energy density of SD is an important
quantity since it impacts the steady state electron and positron
distribution through cooling; a high energy density of SD photons in
a cluster suppresses the gamma-ray flux from IC upscattered
photons. We derive the energy density from starlight and dust in a
galaxy cluster through
\begin{eqnarray}
\label{eq:u_SD_r}
u_\sd(r) &=& \int \dd \eph \frac{\dd^2 N_\ph}{\dd \eph \dd V}\,\eph
=\int \dd \eph \frac{u_\sd(\eph, r)}{\eph}
\nonumber \\
&=&  j(r)  \int \dd \eph \sum_i 
N_i(\mvir) \frac{u_i^\gal(\eph)}{\eph}\,, \nonumber \\
\end{eqnarray}
where we have used that specific energy density of the SD light given
by Eq.~(\ref{eq:u_SD_er}). Note that we compare the energy density
$u_\sd(r)$ to other radiation background fields in
Fig.~\ref{fig:SD_Edens} for a galaxy cluster with the mass
$\mvir=6\times10^{14}\msun$. We find that inside a few percent of of
$\rvir$ the total energy density is dominated by the light from SD,
while the CMB is dominating elsewhere. This implies that if the boost
from dark matter (DM) substructures is significant then the overlap of
light from SD with the electron and positron distribution that trace
the substructures is small. Hence the resulting flux from inverse
Compton (IC) upscattered SD photons is suppressed compared to the IC
upscattered CMB photons.

@@

\section{Source functions of different DM models}
We use {\sc DarkSUSY} to derive the spectra for the continuum emission
from our four DM benchmark (BM) models. These spectra are shown in the
left panel of Fig.~\ref{fig:q_DM}. We choose to fit the source
function using the following functional form
\begin{eqnarray}
q_\rmn{BM} (\eg,r)&=&\left[\frac{\rho(r)}{10^{-29}\rmn{g}\,\cm^{-3}}\right]^2
\times\frac{a_1}{\egt}\times\theta(\eg-m_\chi\,c^2)\nonumber\\
&\times&\left[\frac{\exp\left(-\egt^{-a_2}\right)}{1+\exp\left(\egt^{a_3}\right)}
+\left(\frac{\egt}{a_4}\right)^{4.5}\right]\,, \nonumber\\
&&\rmn{where}\quad \egt = \frac{\eg}{\gev}\,,
 \label{eq:bm_cont}
\end{eqnarray}
where the model specific parameters are given in
Tab.~\ref{tab:bm_cont}.

\begin{table}[h]
\begin{tabular}{c c c c c }
\hline
\hline
 BM model & $a_1$ & $a_2$  & $a_3$ & $a_4$ \\
          & $[\gev^{-1}\,\cm^{-2}\,\rmn{s}^{-1}]$ & & & \\
 \hline
$\Ip$ & $2.0\times10^{-4}$ & $0.36$ & $0.51$ & $701$  \\
$\Jp$ & $4.1\times10^{-6}$ & $0.40$ & $0.43$ & $1178$ \\
$\Kp$ & $1.7\times10^{-4}$ & $0.42$ & $0.37$ & $-$    \\
$\Js$ & $2.1\times10^{-6}$ & $0.34$ & $0.51$ & $490$  \\
\hline
\hline
\end{tabular}
\begin{quote}
The best fit parameters to the continuum emission from our
supersymmetric BM models.
 \label{tab:bm_cont}
  \end{quote}
\end{table}

Again we use {\sc DarkSUSY} to derive the number of electrons and
positrons resulting from each annihilating pair. This spectra is shown
in the right panel of Fig.~\ref{fig:q_DM}. The electron and positron
yield resulting from our supersymmetric BM model is fitted using
\begin{eqnarray}
N_{\e,\rmn{BM}}(>\ee) &=& \frac{b_1}{\eet^{b_2}}\times
\frac{\left[1+\left(\frac{\eet}{b_6}\right)^{b_7}\right]^{b_8}}
{\left[1+\left(\frac{\eet}{b_3}\right)^{b_4}\right]^{b_5}}
\times\theta(\ee-m_\chi\,c^2)\,,\nonumber\\
& &\rmn{where}\quad\eet=\frac{\ee}{\gev}\,.
\label{eq:bm_elec}
\end{eqnarray}
The best fit parameters are given in Tab.~\ref{tab:bm_elec}.
\begin{table}[h]
\begin{tabular}{c c c c c c c c c}
\hline
\hline
 BM model & $b_1$ & $b_2$ & $b_3$ & $b_4$ & $b_5$ & $b_6$ & $b_7$ & $b_8$ \\
 \hline
$\Ip$ & $28$ & $0.06$ & $1.8$ & $0.9$  & $2.0$ & $60$  & $5$ & $-1$     \\
$\Jp$ & $34$ & $0.08$ & $5.5$ & $0.8$  & $2.6$ & $250$ & $5$ & $-1$     \\
$\Kp$ & $44$ & $0.05$ & $1.7$ & $0.85$ & $2.1$ & $10$  & $2$ & $0.62$   \\
$\Js$ & $30$ & $0.14$ & $5.0$ & $1.0$  & $1.5$ & $300$ & $5$ & $-1$     \\
\hline
\hline
\end{tabular}
\begin{quote}
The best fit parameters to the electron and positron yield from the DM
BM models above the electron energy $\ee$.
 \label{tab:bm_elec}
  \end{quote}
\end{table}

Once again, we use {\sc DarkSUSY} to derive the electron and positron
yield resulting from LP DM annihilating into decaying muons. We find
the following best fit:
\begin{eqnarray}
N_{\e,\rmn{LP}\mu}(>\ee) &=& \exp\left(c_1+c_2\,y+c_3\,y^2\right)
\,,\nonumber\\
\rmn{where}\quad y&=&\,\log\left(1-x\right)\,,\quad 
x=\frac{\ee}{m_\chi\,c^2}\,\quad\rmn{and}\nonumber\\
c_1&=&0.719,\,c_2=2.27,\,c_3=0.200\,.
\label{eq:lp_elec}
\end{eqnarray}
It should be noted that the LP model we are using has a branching
factor branching ratio of ($1/2,\,1/2$) into
($\mu^+/\mu^-,\,\e^+/\e^-$). Hence the resulting electron and positron
yield for the LP model is given by 
\begin{equation}
N_{\e,\rmn{LP}}(>\ee) =
1+N_{\e,\rmn{LP}\mu}(>\ee)/2\,.
\end{equation}

>>>>

\section{Detailed modelling of starlight and dust in clusters}
\label{sect:SD}
In this section we present the details of our model of stars and dust
(SD) in clusters of galaxies. In the first part we use the measured
spectral distribution of light from SD in a typical galaxy to derive a
simple spectral model for SD in galaxy clusters. This is done by
renormalizing the galactic SD light to match the observed luminosity
from a cluster in the far infrared (IR) and ultraviolet (UV) bands. In
the second part we derive the spatial distribution of the SD light
from a stacked sample of clusters.

\subsection{Energy spectrum}
The light from SD in clusters is dominated by the emission from
galaxies. This radiation is expected to leak out into the much larger
volume of the intra-cluster medium (ICM), hence we approximate the
spectral distribution of the intra-cluster light (ICL) with the light
from galaxies. We use the far IR to UV spectrum derived in reference
\cite{2006ApJ...648L..29P} for a Milky Way (MW) type galaxy to
characterize the spectral distribution of SD in a cluster. Note that
we only keep the spectral shape, and renormalize the spectral
distribution using the luminosity from SD in clusters.

In Fig.~\ref{fig:SD_spectra} we show the spectral distribution of the
light from SD together with the cosmic microwave background (CMB)
black-body distribution in a $6\times10^{14}\,\msun$ galaxy
cluster. We fit the spectral shape of SD individually; the dust that
peaks at about $10^{-2}\,\ev$ is fitted using a double power-law,
while the broader spectral distribution of the stars that peaks at
about $1\,\ev$ is fitted using a tipple power-law. The best fit to the
galactic spectrum is given by:
\begin{eqnarray}
  u_\stars^\gal(\eph) &=& \frac{23\,\rmn{eV}}{\rmn{cm}^3} 
  \left(\frac{1.23\,\rmn{eV}}{\eph}\right)^{1.9} \nonumber \\
  &\times&\left[1+\left(\frac{2.04\,\rmn{eV}}{\eph}\right)^{20}\right]
  ^{-\frac{1.9}{20}}\nonumber \\
  &\times& \left[1+\left(\frac{0.78\,\rmn{eV}}{\eph}\right)^{20}\right]^{-\frac{2.6}{20}}\,, \\
  u_\dust^\gal(\eph) &=& 
  \frac{40\,\rmn{eV}}{\rmn{cm}^3} 
  \left(\frac{0.0144\,\rmn{eV}}{\eph}\right)^{4.9}\nonumber \\
  &\times& \left[1+\left(\frac{0.0144\,\rmn{eV}}{\eph}\right)^{1.9}\right]^{-4.9}\,.
\end{eqnarray}
The specific energy density of the light from SD in a cluster can now
be derived from
\begin{eqnarray} 
u_\sd(\eph, r) &=& j(r)\,u_\sd(\eph)\,,
\label{eq:u_SD_er}
\end{eqnarray}
where the spectral distribution of SD is given by
\begin{eqnarray}
  u_\sd(\eph) &\equiv& \eph^2\frac{\dd^2 N_\ph}{\dd \eph \dd V}
  =  \sum_i N_i(\mvir) u_i^\gal(\eph)\,,\nonumber \\ 
\rmn{for}\quad i&=&\{\rmn{stars,dust}\}\,.
\end{eqnarray}
Here $\eph$ denotes the energy of the SD photons, $j(r)$ describes the
unitless spatial distribution of SD (derived in
sect.~\ref{sect:SD_radial}), and $N_i(\mvir)$ denotes the mass
dependent normalization for the SD. This function is determined using
the total energy of the light from SD within $\rvir$ which is
represented by $E_{i,\vir}$. We relate this energy to the luminosity
$L_i$ in each respective wavelength band (optical and IR light for the
dust, and UV light for the stars) using
\begin{eqnarray} 
  E_{i,\vir} &=& L_i \frac{\rvir}{c} \nonumber \\
  &=&N_i(\mvir)\int_{\rvir} \int_i \,j(r) 
  \frac{u_i^\gal(\eph)}{\eph}\,\dd V\dd \eph\,,\nonumber \\
 \rmn{for}\quad i&=&\{\rmn{stars,dust}\}\,.
\label{eq:E_SD}
\end{eqnarray}
Here we approximate the total energy of the photons within $\rvir$
with the SD luminosity multiplied the typical timescale it takes for a
photon to propagate through the cluster (we assume that the cluster is
optically thin).

We derive the {\em luminosity from the stars} using
\begin{eqnarray}
L_\stars(M_{200\rmn{av}})&=&10^{\frac{m_\odot-m_\rmn{cl}-2.5\log\left(1+z\right)}{2.5}}
\left(\frac{D_\clu}{D_\odot}\right)^2 L_\odot\nonumber\\
&\approx& 5.3\times10^{44}\,\rmn{erg}\,\rmn{s}^{-1}\,,
\label{eq:L_stars}
\end{eqnarray}
where the average apparent magnitude of a cluster at $z\approx 0.25$
in the $(r+i)$-band is given by $m_\rmn{cl}\approx 15.5$
\cite{2005MNRAS.358..949Z}. This magnitude is derived from stacked
clusters observed by the SDSS with the average mass
$M_{200\rmn{av}}=4.0\times10^{13}\msun$. We also use an apparent
magnitude of the sun $m_\odot\approx -27.7$, distance to the sun
$D_\odot=4.85\times10^{-6}\,\rmn{pc}$, luminosity of the sun
$L_\odot=1.17\times10^{33}\,\rmn{erg\,s}^{-1}$, and distance to the
cluster $D_\clu=1.26\times10^9\,\rmn{pc}$. Furthermore, to account for
different cluster masses, we attach a simple mass scaling to the
luminosity from the stars in Eq.~(\ref{eq:L_stars}). Here we assume
that the total starlight has the same halo mass scaling as the
brightest cluster galaxy (BCG) \cite{2010ApJ...713.1037H}, which is a
reasonable assumption since a large contribution of the total
starlight in a cluster comes from the BCG. The starlight luminosity as
a function of mass is given by
\begin{equation}
L_\stars(\mvir)=L_\stars(M_{200\rmn{av}})\,
\left(\frac{\mvir}{M_{200\rmn{av}}}\right)^{0.18}\,.
\label{eq:L_stars_m}
\end{equation} 
We can now fix the unitless normalization constant for the stars in
Eq.~(\ref{eq:E_SD}) by integrating over the cluster volume and
spectral distribution of the stars:
\begin{equation}
 N_\stars(\mvir) = 
\left(\frac{\mvir}{M_{200\rmn{av}}}\right)^{0.18}\,
\frac{6.0\,\rvir\times10^{-9}\,\kpc^2}{\int_{\rvir} j(r) \,\dd V}\,.
\label{eq:N_stars}
\end{equation}

@@

\subsection{Radial profiles}
\label{sect:SD_radial}
Our goal in this section is to derive a simple spatial model for the
distribution of the light from SD in galaxy clusters. For this reason
we use stacked cluster observations by Sloan in the $r$ and $i$-band
that trace the stars and for simplicity we assume that the dust trace
the stars in the clusters {\bf Christoph, why is this a valid
  approximation?}. The stacked surface brightness profiles in
\cite{2005MNRAS.358..949Z} are measured at redshift $z \approx 0.25$
with an average mass of the clusters of $4.0\times10^{13}\,\msun$. The
profiles are build up out of three components; diffuse ICL, galaxies,
and the BCG in the center of clusters. However, the total volume
overlap of the inter-stellar medium with a cluster is very small
compared to the total overlap of the ICL- and BCG with a cluster, thus
the relative contribution of the galaxies to the IC emission is
suppressed (c.f. Eq.~\ref{eq:SD_overlap}). We remove this bias in our
benchmark model for SD by cutting out the galaxies which roughly
corresponds to $70$ percent of the total SD light. In
Fig.~\ref{fig:SD_spatial} we show the SDSS stacked brightness profiles
as well as the fitted profiles\footnote{The measured brightness is
  converted into units of $\rmn{erg}\,\rmn{s}^{-1}$ using
  \cite{2010...book} $S(\rmn{mag}\,''^{-2}) =
  \mathcal{M}_\odot+21.6-2.5\log_{10}[S(L_\odot\,\rmn{pc}^{-2})]$,
  where the sun's absolute magnitude in the $r$-band is given by
  $\mathcal{M}_\odot=27.1$ \cite{1998gaas.book.....B} and the
  luminosity of the sun by $L_\odot=3.85\times10^{33}\,
  \rmn{erg}\,\rmn{s}^{-1}$.}, where both the data and fits are
normalized with the central surface brightness of the total SD
component ($S_\sd^\rmn{tot}(0)$). Our spatial benchmark model is shown
with the solid black line, and include the contribution from the ICL
and BCG.

@@

Instead of modelling the surface brightness with a de Vaucouleur
profile with the addition of a power-law, we use a double beta profile
model for simplicity of deprojection. It is given by
\begin{equation}
\tilde{S}_\sd (r_\bot) = \frac{S_\sd (r_\bot)}{S_\rmn{SD}^\rmn{tot}(0)} = 
\sum_{i=1}^2 \,\tilde{S}_i\, 
\left[1 + \left( \frac{r_\bot}{r_{\mathrm{c}_i}}\right)^2\right]
^{-3\beta_i + 1/2}\,,
\label{double_beta}
\end{equation}
where we found the following best fitted normalized central brightness
($\tilde{S}_i$), core radius ($r_{\mathrm{c}_i}$), and slope
($\beta_i$) for the total model:
\begin{align}
&\tilde{S}_1^\rmn{tot} = 1.0\,,
&&r_{\mathrm{c}_1}^\rmn{tot} = 1.8\,\rmn{kpc}\,,
&&&\beta_1^\rmn{tot} = 0.45\, \nonumber\\
&\tilde{S}_2^\rmn{tot} = 2.3\times10^{-3}\,,
&&r_{\mathrm{c}_2}^\rmn{tot} = 0.19\,\rvir\,,
&&&\beta_2^\rmn{tot} = 0.44\,.\nonumber\\
& && &&&
\label{fit_spatial_IR}
\end{align}
For the smooth model, which is also our benchmark SD model, we find:
\begin{align}
&\tilde{S}_1^\rmn{sm}=0.47,\,
&&r_{\mathrm{c}_1}^\rmn{sm}=3.9\,\rmn{kpc}\,,
&&&\beta_1^\rmn{sm}=0.53 \,\nonumber\\
&\tilde{S}_2^\rmn{sm}=8.3\times10^{-4}\,,
&&r_{\mathrm{c}_2}^\rmn{sm}=0.19\,\rvir\,,
&&&\beta_2^\rmn{sm}=0.54\,.\nonumber\\
& && &&&
\label{fit_spatial_IR_sm}
\end{align}
The three dimensional spatial profile is derived by deprojecting the
surface brightness in Eq.~(\ref{double_beta}) (see
e.g. \cite{2004A&A...413...17P} for details about the deprojection):
\begin{eqnarray}
  j(r)  = \sum_{i=1}^2 \frac{\tilde{S}_i}{2\pi\,r_{\mathrm{c}_i}}\,
  \frac{6 \beta_i - 1}{\left(1 + r^2/r^2_{\mathrm{c}_i}\right)^{3\beta_i}}\,
  \mathcal{B}\left(\frac{1}{2},3\beta_i\right)\,,\nonumber\\
\end{eqnarray}
where $\mathcal{B}(a,b)$ denotes the beta-function
\cite{1965hmfw.book.....A}.

The spatial distribution of the energy density of SD is an important
quantity since it impacts the steady state electron and positron
distribution through cooling; a high energy density of SD photons in
a cluster suppresses the gamma-ray flux from IC upscattered
photons. We derive the energy density from starlight and dust in a
galaxy cluster through
\begin{eqnarray}
\label{eq:u_SD_r}
u_\sd(r) &=& \int \dd \eph \frac{\dd^2 N_\ph}{\dd \eph \dd V}\,\eph
=\int \dd \eph \frac{u_\sd(\eph, r)}{\eph}
\nonumber \\
&=&  j(r)  \int \dd \eph \sum_i 
N_i(\mvir) \frac{u_i^\gal(\eph)}{\eph}\,, \nonumber \\
\end{eqnarray}
where we have used the specific energy density of the light from SD
given by Eq.~(\ref{eq:u_SD_er}). Note that we compare the energy
density $u_\sd(r)$ to other radiation background fields in
Fig.~\ref{fig:SD_Edens} for a galaxy cluster with the mass
$\mvir=6\times10^{14}\msun$. We find that inside a few percent of of
$\rvir$ the total energy density is dominated by the light from SD,
while the CMB is dominating elsewhere. This implies that if the boost
from dark matter (DM) substructures is significant, then the overlap of
light from SD with the electron and positron distribution that trace
the substructures is small. Hence the resulting flux from inverse
Compton (IC) upscattered SD photons is suppressed compared to the IC
upscattered CMB photons.

@@

\section{Source functions of different DM models}
We use {\sc DarkSUSY} to derive the spectral distribution of the
continuum emission from our four DM benchmark (BM) models. These
spectra are shown in the left panel of Fig.~\ref{fig:q_DM}. We choose
to fit the source function using the following functional form
\begin{eqnarray}
q_\rmn{BM} (\eg,r)&=&\left[\frac{\rho(r)}{10^{-29}\rmn{g}\,\cm^{-3}}\right]^2
\times\frac{a_1}{\egt}\times\theta(\eg-m_\chi\,c^2)\nonumber\\
&\times&\left[\frac{\exp\left(-\egt^{-a_2}\right)}{1+\exp\left(\egt^{a_3}\right)}
+\left(\frac{\egt}{a_4}\right)^{4.5}\right]\,, \nonumber\\
&&\rmn{where}\quad \egt = \frac{\eg}{\gev}\,,
 \label{eq:bm_cont}
\end{eqnarray}
where the model specific parameters are given in
Tab.~\ref{tab:bm_cont}.

\begin{table}[h]
\begin{tabular}{c c c c c }
\hline
\hline
 BM model & $a_1$ & $a_2$  & $a_3$ & $a_4$ \\
          & $[\gev^{-1}\,\cm^{-2}\,\rmn{s}^{-1}]$ & & & \\
 \hline
$\Ip$ & $2.0\times10^{-4}$ & $0.36$ & $0.51$ & $701$  \\
$\Jp$ & $4.1\times10^{-6}$ & $0.40$ & $0.43$ & $1178$ \\
$\Kp$ & $1.7\times10^{-4}$ & $0.42$ & $0.37$ & $-$    \\
$\Js$ & $2.1\times10^{-6}$ & $0.34$ & $0.51$ & $490$  \\
\hline
\hline
\end{tabular}
\caption{The best fit parameters to the continuum emission from our
  supersymmetric benchmark models.
 \label{tab:bm_cont}}
\end{table}

Once again we use {\sc DarkSUSY} to derive the number of electrons and
positrons resulting from each annihilation. This spectra is shown in
the right panel of Fig.~\ref{fig:q_DM}. The electron and positron
yield resulting from our supersymmetric BM model is fitted using
\begin{eqnarray}
N_{\e,\rmn{BM}}(>\ee) &=& \frac{b_1}{\eet^{b_2}}\times
\frac{\left[1+\left(\frac{\eet}{b_6}\right)^{b_7}\right]^{b_8}}
{\left[1+\left(\frac{\eet}{b_3}\right)^{b_4}\right]^{b_5}}
\times\theta(\ee-m_\chi\,c^2)\,,\nonumber\\
& &\rmn{where}\quad\eet=\frac{\ee}{\gev}\,.
\label{eq:bm_elec}
\end{eqnarray}
The best fit parameters are given in Tab.~\ref{tab:bm_elec}.
\begin{table}[h]
\begin{tabular}{c c c c c c c c c}
\hline
\hline
 BM model & $b_1$ & $b_2$ & $b_3$ & $b_4$ & $b_5$ & $b_6$ & $b_7$ & $b_8$ \\
 \hline
$\Ip$ & $28$ & $0.06$ & $1.8$ & $0.9$  & $2.0$ & $60$  & $5$ & $-1$     \\
$\Jp$ & $34$ & $0.08$ & $5.5$ & $0.8$  & $2.6$ & $250$ & $5$ & $-1$     \\
$\Kp$ & $44$ & $0.05$ & $1.7$ & $0.85$ & $2.1$ & $10$  & $2$ & $0.62$   \\
$\Js$ & $30$ & $0.14$ & $5.0$ & $1.0$  & $1.5$ & $300$ & $5$ & $-1$     \\
\hline
\hline
\end{tabular}
\caption{The best fit parameters to the electron and positron yield
  above the electron energy $\ee$ from the dark matter benchmark
  models.
 \label{tab:bm_elec}}
\end{table}

Finally, we use {\sc DarkSUSY} to derive the electron and positron
yield resulting from LP DM annihilating into decaying muons. We find
the following best fit:
\begin{eqnarray}
N_{\e,\rmn{LP}\mu}(>\ee) &=& \exp\left(c_1+c_2\,y+c_3\,y^2\right)
\,,\nonumber\\
\rmn{where}\quad y&=&\,\log\left(1-x\right)\,,\quad 
x=\frac{\ee}{m_\chi\,c^2}\,\quad\rmn{and}\nonumber\\
c_1&=&0.719,\,c_2=2.27,\,c_3=0.200\,.
\label{eq:lp_elec}
\end{eqnarray}
It should be noted that the LP model that we are using in this work
has a branching factor branching ratio of ($1/2,\,1/2$) into
($\mu^+/\mu^-,\,\e^+/\e^-$). Hence the resulting electron and positron
yield for the LP model is given by
\begin{equation}
N_{\e,\rmn{LP}}(>\ee) =
1+N_{\e,\rmn{LP}\mu}(>\ee)/2\,.
\end{equation}
